---
title: "Data import"
output: html_document
---
```{r setuo, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)

```
This document will show how to import data.

## Import the FAS_Litters CSV and clean the name

```{r}
litters_df = read_csv("data/FAS_litters.csv")

litters_df =janitor::clean_names(litters_df)
```
Look at the data
```{r}
litters_df
head(litters_df)
tail(litters_df, 10)
``` 
```{r echo=FALSE}
view(litters_df)
```

#exerceise of import and clean name
```{r}
pups_df = read_csv(file = "data/FAS_pups.csv")

pups_df = janitor::clean_names(pups_df)

pups_df
```
##look at the read_csv options(skipping lines)

```{r}
litters_df = read.csv(
  file = "data/FAS_litters.csv",
  header = FALSE, 
  skip = 1     #first line skip, if =2 is the first two line skip    
)

```
## what about missing data
```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA","",".")
  )

litters_df = janitor::clean_names(litters_df)

pull(litters_df, gd0_weight)
```
##what if we code'group' as a factor variable?
```{r}
litters_df = 
  read_csv(
  
    file = "data/FAS_litters.csv",
    na = c("NA", "", "."),
  
    col_types = cols(
      Group = col_factor()  
    )
  )

```
##import an excel file

Import MLB 2011 summary data.
```{r}
mlb_df = read_excel("data/mlb11.xlsx", sheet ="mlb11")
```

##Import SAS data
```{r}

pulse_df = read_sas("data/public_pulse_data.sas7bdat")


```
##never use read.csv()


Data manipulate

Import two datasets going to manipulate

```{r}
litters_df=
  read_csv("data/FAS_litters.csv", na = c("NA", "", ".") )

itters_df = janitor::clean_names(litters_df)


pups_df = 
    read_csv("./data/FAS_pups.csv", na = c("NA", "."))

pups_df = janitor::clean_names(pups_df)

```

##Select

select the function

```{r}
select(litters_df,
       litter_number, gd0_weight, pups_born_alive)
```

```{r}
select(litters_df, group:gd18_weight)
```
```{r}
select(litters_df, -(group:gd18_weight))

select(litters_df, starts_with("gd"))

select(litters_df, contains("pups"))

```

rename
```{r}

select(litters_df, Group = group)
rename(litters_df, Group = group)
select(litters_df, litter_number, gd_0weight, everything())

relocate(litters_df, litter_number, gd_0weight)

```
exercise about select

```{r}
select(pups_df, litter_number, sex, pd_ears)
```


## filter

```{r}
filter(litters_df, gd_of_birth ==20)

filter(litters_df, gd_of_birth ==20)

```
```{r}
filter(litters_df, gd_of_alive >5)

filter(litters_df, gd_of_alive !=5) #not equal to 5
```

```{r}
filter(litters_df, group =="Low8")

filter(litters_df, group %in% c("Low8", "Low8"))

filter(litters_df, group %in% c("Low8") | pups_born_alive == 8) #means low8 or xxx==8 is satisfied with the standard;逻辑或运算符 | 表示：如果 group 列值等于 "Low8"，或者 pups_born_alive 列的值等于 8，就保留该行
```

```{r}
drop_na(litters_df) #删除任何列中包含缺失值的行
drop_na(litters_df, gd0_weight) #只删除 gd0_weight 列中有缺失值的行
```

assesment

```{r}
filter(pups_df, sex == 1)#Filter to include only pups with sex 1


filter(pups_df, sex == 2, pd_walk < 11) #Filter to include only pups with PD walk less than 11 and sex 2
```

## mutate

```{r}
mutate(litters_df,
  wt_gain = gd18_weight - gd0_weight,  # 创建一个名为 wt_gain 的新列，表示差值
  gd18_weight_squared = gd18_weight^2, # 创建一个名为 gd18_weight_squared 的新列，表示 gd18_weight 的平方
  group = str_to_lower(group)           # 将 group 列的值转换为小写
)

```

## arange
```{r}
arrange(litters_df, group)

arrange(litters_df, desc(pups_born_alive))
```

```{r}
head(arrange(litters_df, group, pups_born_alive), 10) #从 litters_df 数据框中按照指定列排序后，提取前 10 行
```

## PIPING!!!!!!
do not do this
```{r}
litters_df_raw = 
    read_csv("./data/FAS_litters.csv", na = c("NA", ".", ""))
litters_df_clean_names = janitor::clean_names(litters_df_raw)
litters_df_selected_cols = select(litters_df_clean_names, -pups_survive) #Many steps
```

do not do this
```{r}
litters_df_clean = 
  drop_na(
    mutate(
      select(
        janitor::clean_names(
          read_csv("./data/FAS_litters.csv", na = c("NA", ".", ""))
          ), 
      -pups_survive
      ),
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)
    ),
  wt_gain
  )

```

Do this!!!!

```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv", na = c("NA", ".", "")) %>% 
  
  janitor::clean_names() |> 
  
  select(-pups_survive) |> 
  
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) |> 
  
  drop_na(wt_gain)

write_csv(litters_df, "data/cleaned_fas_litters.csv")
```
```{r}
litters_df |>
  lm(wt_gain ~ pups_born_alive, data = _) |>
  broom::tidy()

#使用 litters_df 数据框，构建一个线性回归模型，解释 wt_gain（体重增加）如何受 pups_born_alive（出生存活的小鼠数）的影响。第二步：将线性回归模型的输出传递给 broom::tidy()，生成一个包含回归结果的数据框，其中包含回归系数、标准误差、t 值和 p 值等统计信息。
```

Data export






























